## Homework 4

<!--
Remember you can use the example4.Rmd or other files to find examples
of the codes you'll need to use to complete this assignment. I strongly
urge you to copy and paste relevant code that you know works from my
examples or instructions and then edit that code so you're working
with the appropriate variables for the homework.
-->

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->

<!--
Read in the homework4.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
--> The first thing to do is read in the data
```{r}
dat<-read.csv("homework4.csv")
```



<!--
Check the structure of the data.
--> The next step is to check the structure of the data

```{r}
str(dat)
```

We see that dat is a data frame that has two variables: Agreement (whether or not the subjects agree with a certain statement), Condition (whether they are in the experimental or control group).
<!--
Create a table that shows how many people fell into each category.
--> I create a table to show how many people fell into each category
```{r}
table(dat$condition, dat$agreement)
```



<!--
Perform a chi-square test to determine whether or not the experimental
conditions and agreement with a statement are independent or related.
Also, interpret the results of your test and provide a a statistical
statement that supports your interpretation. As there's no way to
insert the chi-square symbol, you can just use chi-square instead
(see example4.Rmd).
--> In order to determine if the data support a relationship between Condition and Agreement, I performed a chi-square test. 
```{r}
chisq.test(dat$condition,dat$agreement)
```
According to the results of the chi-square test, condition and agreement are independent of one another Chi-square(1) = 1.3, p = .26
